NumPy---Array Programming in Python

### 2020-01-14 body.tex
Note to add references to Parquet, Arrow, Zarr, and Ruby's Numo
Combine figures

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{static/sketches/array-concepts}
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated meta-data fields.
    \textbf{b,} Indexing an array with various types of arguments.
    \textbf{c,} Broadcasting in scalar addition, and in the division of two-dimensional arrays.
    \textbf{d,} Indexing with arrays, which broadcast the indexing arguments before performing the lookup.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

Indexing an array returns single elements, sub-arrays, or elements that satisfy
a specific condition (Fig.~\ref{fig:array-concepts}b).
Arrays can even be indexed using other arrays (Fig.~\ref{fig:array-concepts}d).
Wherever possible, indexing that retrieves a subarray returns a {\em view} on
the original array, such that data is shared between the two arrays.
This provides a powerful way to operate on subsets of array data, while
limiting memory usage.

A powerful array feature that affects the way NumPy indexes and combines
arrays, is {\em broadcasting}.
When performing an operation (such as addition) on two arrays, one intuitively
expect that these arrays should be required to have the same shape.
However, through broadcasting, NumPy allows the dimensions to differ, while
still producing results that appeal to intuition.
A trivial example is the addition of a scalar value to an array, but it also
generalizes to more complex examples such as scaling each column of an array,
or generating a grid of coordinates.
In broadcasting, one or both arrays are virtually duplicated (that is, without
copying any data in memory), so that the shapes of the operands match
(Fig.~\ref{fig:array-concepts}c).
Broadcasting is also applied whenever an array is indexed using arrays of
indices (Fig.~\ref{fig:array-concepts}d).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{static/sketches/reductions}
  \caption{\textbf{Reductions.} Reduction operations act along one or more axes. In this example, a three-dimensional array is shown to be summed along various single axes to produce two-dimensional results, or along two axes consecutively to produce a one-dimensional result. \fixme{May want to fix to have right-hand axis system.}
   }
  \label{fig:reductions}
\end{figure}

@@

Many of these functions, such as summation, support reductions: aggregating
results across over or more dimension of the array.
For example, summing a $D$-dimensional array over $p$ axes results in a
$D-p$-dimensional array (Fig.~\ref{fig:reductions}).

@@

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=.9\textwidth]{static/sketches/array-concepts}   
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated meta-data fields.
    \textbf{b,} Indexing an array with various types of arguments.
    \textbf{c,} Indexing with arrays, which broadcast the indexing arguments before performing the lookup.
    \textbf{d,} Broadcasting in scalar addition, and in the division of two-dimensional arrays.
    \textbf{e,} Reduction operations act along one or more axes. In this
    example, a three-dimensional array is shown to be summed along various single
    axes to produce two-dimensional results, or along two axes consecutively to
    produce a one-dimensional result.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

Indexing an array returns single elements, sub-arrays, or elements that satisfy
a specific condition (Fig.~\ref{fig:array-concepts}b).
Arrays can even be indexed using other arrays (Fig.~\ref{fig:array-concepts}c).
Wherever possible, indexing that retrieves a subarray returns a {\em view} on
the original array, such that data is shared between the two arrays.
This provides a powerful way to operate on subsets of array data, while
limiting memory usage.

A powerful array feature that affects the way NumPy indexes and combines
arrays, is {\em broadcasting}.
When performing an operation (such as addition) on two arrays, one intuitively
expect that these arrays should be required to have the same shape.
However, through broadcasting, NumPy allows the dimensions to differ, while
still producing results that appeal to intuition.
A trivial example is the addition of a scalar value to an array, but it also
generalizes to more complex examples such as scaling each column of an array,
or generating a grid of coordinates.
In broadcasting, one or both arrays are virtually duplicated (that is, without
copying any data in memory), so that the shapes of the operands match
(Fig.~\ref{fig:array-concepts}d).
Broadcasting is also applied whenever an array is indexed using arrays of
indices (Fig.~\ref{fig:array-concepts}c).

@@

Many of these functions, such as summation, support reductions: aggregating
results across over or more dimension of the array.
For example, summing a $D$-dimensional array over $p$ axes results in a
$D-p$-dimensional array (Fig.~\ref{fig:array-concepts}e).

@@

\fixme{Add references somewhere to Zarr, https://arrow.apache.org/, https://parquet.apache.org/, https://github.com/ruby-numo/numo-narray/wiki/Comparison-with-Numpy}
